Quality Definitions of Qualities

### 2021-10-15 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
added mathematical definition of abstraction comment - fixed error
Updates to the definition of the quality of usability

  \item[Consistency] The final list of definitions should be consistent in the
  terminology used.  For instance, we do not use synonyms; we use
  the same word or phrase to mean the same concept throughout the final list of
  definitions.  Specific decisions on terminology are as follows:
  \begin{itemize}
    \item To represent the idea of cost, effort, time, resources, and
    efficiency, we say \emph{effort}.
    \item To represent ``the degree to which'' or the ``the extent to which'',
    we will say \emph{the extent to which}.
    \item \wss{Fill in additional consistency decisions, as we go through
    definitions.}
  \end{itemize}

@@

\begin{defn}
  ``The effort required to learn, operate, prepare input, and interpret output
  of a program'' \citep{McCallEtAl1977}.
\end{defn}

\begin{defn}
  ``A software is usable - or user friendly - if its human users find it easy
  to use'' \citep{ghezzi1991fundamentals}. \wss{switch to 2003 ref?}
\end{defn}

\begin{defn}
  The extent to which a product can be used by specified users to achieve
  specified goals with effectiveness, efficiency, and satisfaction in a
  specified context of use. \cite{ISO16982:2002} and \cite{ISO9241:11}
\end{defn}

Nielsen and (separately) Schneidermann have defined usability as part of
usefulness and is composed of:
\begin{defn}
\begin{itemize}
\item Learnability: How easy is it for users to accomplish basic tasks the first
  time they encounter the design?
\item Efficiency: Once users have learned the design, how quickly can they
  perform tasks?
\item Memorability: When users return to the design after a period of not using
  it, how easily can they re-establish proficiency?
\item Errors: How many errors do users make, how severe are these errors, and
  how easily can they recover from the errors?
\item Satisfaction: How pleasant is it to use the design?
\end{itemize}
In that context, it makes sense to separate \emph{usefulness} into
\emph{usability} (purely an interface concern) and \emph{utility} (in the
economics sense of the word).
\cite{NielsenNorman} %2012
\end{defn}

\begin{defn}
Desirable outcomes of usability:
\begin{enumerate}
\item It should be easy for the user to become familiar with and competent in
using the user interface during the first contact with the website. For
example, if a travel agent's website is a well-designed one, the user should
be able to move through the sequence of actions to book a ticket quickly.
\item It should be easy for users to achieve their objective through using the
  website. If a user has the goal of booking a flight, a good design will guide
  him/her through the easiest process to purchase that ticket.
\item It should be easy to recall the user interface and how to use it on
  subsequent visits. So, a good design on the travel agent's site means the user
  should learn from the first time and book a second ticket just as easily.
\end{enumerate}
From the Interaction Design Foundation
\url{https://www.interaction-design.org/literature/topics/usability}.
\end{defn}

One core reference, for definitions and metrics, is
\citet{bevan1995measuring}.

\begin{defn}
  The capability of the software to be understood, learned, used and liked by
  the user, when used under specified conditions \cite{ISO9126}.
\end{defn}

\begin{mybox}
\subsubsection*{Usability}
The extent to which a product can be used by specified users to achieve
specified goals with effectiveness, efficiency, and satisfaction in a specified
context of use. \cite{ISO16982:2002} and \cite{ISO9241:11}
\end{mybox}

@@

There has been a lot of thought put behind the ISO definition, from experts. The
standard has been revisited, but the definition has not changed, which means
that it has stood the test of time.  The different reasons for the proposed
definition are as follows:

@@

We are not exploring a mathematical definition of abstraction at this time. \cite{ma1991types} discuss type-based abstraction captured in mathematical theory.

>>>>

  \item[Consistency] The final list of definitions should be consistent in the
  terminology used.  For instance, we do not use synonyms; we use
  the same word or phrase to mean the same concept throughout the final list of
  definitions.  Specific decisions on terminology are as follows:
  \begin{itemize}
    \item To represent the idea of cost, effort, time, resources, and
    efficiency, we say \emph{effort}.
    \item To represent ``the degree to which'' or the ``the extent to which'',
    we will say \emph{the extent to which}.
    \item Definitions use phrases like ``goals'' and ``requirements''.  We will
    use the term ``requirements'', since it implies something more concrete than
    a goal.
    \item \wss{Fill in additional consistency decisions, as we go through
    definitions.}
  \end{itemize}

@@

\begin{defn} \label{UsabilityMcCallEtAl1977}
  The effort required to learn, operate, prepare input, and interpret output
  of a program \citep{McCallEtAl1977}.
\end{defn}

\begin{defn} \label{UsabilityISO9126}
  The capability of the software to be understood, learned, used and liked by
  the user, when used under specified conditions. \cite{ISO9126}
\end{defn}

\begin{defn} \label{UsabilityGhezziEtAl2003}
  A software system is usable - or \emph{user friendly} - if its human users
  find it easy to use. \citep[p.\ 22]{GhezziEtAl2003}.
\end{defn}

\begin{defn} \label{UsabilityISO16982:2002}
  The extent to which a product can be used by specified users to achieve
  specified goals with effectiveness, efficiency, and satisfaction in a
  specified context of use. \citep{ISO16982:2002, ISO9241:11}
\end{defn}

\begin{defn} \label{UsabilityNielsenNorman}
Usability as part of usefulness and is composed of:
\begin{itemize}
\item Learnability: How easy is it for users to accomplish basic tasks the first
  time they encounter the design?
\item Efficiency: Once users have learned the design, how quickly can they
  perform tasks?
\item Memorability: When users return to the design after a period of not using
  it, how easily can they re-establish proficiency?
\item Errors: How many errors do users make, how severe are these errors, and
  how easily can they recover from the errors?
\item Satisfaction: How pleasant is it to use the design?
\end{itemize}
In that context, it makes sense to separate \emph{usefulness} into
\emph{usability} (purely an interface concern) and \emph{utility} (in the
economics sense of the word).
\cite{NielsenNorman} %2012
\end{defn}

\begin{defn} \label{UsabilityDesFoundation}
Desirable outcomes of usability:
\begin{enumerate}
\item It should be easy for the user to become familiar with and competent in
using the user interface during the first contact with the website. For
example, if a travel agent's website is a well-designed one, the user should
be able to move through the sequence of actions to book a ticket quickly.
\item It should be easy for users to achieve their objective through using the
  website. If a user has the goal of booking a flight, a good design will guide
  him/her through the easiest process to purchase that ticket.
\item It should be easy to recall the user interface and how to use it on
  subsequent visits. So, a good design on the travel agent's site means the user
  should learn from the first time and book a second ticket just as easily.
\end{enumerate}
From the Interaction Design Foundation
\url{https://www.interaction-design.org/literature/topics/usability}.
\end{defn}

% One core reference, for definitions and metrics, is
% \citet{bevan1995measuring}.

\begin{mybox}
\subsubsection*{Usability}
The extent to which software can be used by specified users to achieve specified
requirements with effectiveness, efficiency, and satisfaction in a specified
context of use. 
\end{mybox}

@@

The selected definition is a modified version of
Definition~\ref{UsabilityISO16982:2002}.  This definition comes from an ISO
standard that has been revisited multiple times, without the definition for
usability changing; therefore, the definition has stood the test of time.
Additional justifications for selecting the proposed definition are as follows:

\begin{itemize}
  \item The selected definition uses our selected phrase of ``the extent to
  which'' for consistency with the other definitions.
  \item The selected definition is less ambiguous compared
  to~\ref{UsabilityMcCallEtAl1977},~\ref{UsabilityISO9126},~\ref{UsabilityGhezziEtAl2003}
  and ~\ref{UsabilityDesFoundation} because the selected definition talks about
  specified users, goals and context of use.
  \item The proposed definition covers the ``learn, operate, prepare input, and
  interpret output'' from Definition~\ref{UsabilityMcCallEtAl1977} and ``Learnability'', Memorability'' and ``Errors'' from Definition~\ref{UsabilityNielsenNorman} by the more
  abstract phrase ``achieve specified requirements''.
  \item We changed ``the product'' from Definition~\ref{UsabilityISO16982:2002}
  to ``software'', since this document covers software qualities.
  \item We changed ``goals'' from Definition~\ref{UsabilityISO16982:2002} to
  ``requirements'' for the goal of being unambiguous and consistent.
\end{itemize}

@@

We are not exploring a mathematical definition of abstraction at this time. \cite{reynolds1983types} discusses type-based abstraction captured in mathematical theory.
